Usage metrics

### 2020-12-11 inst/analysis/usage_metrics.Rmd

---
title: "Usage metrics"
author: "Pierre Donat-Bouillud"
output:
    html_document:
        gallery: false
        toc: true
        toc_depth: 3
        toc_float: true
        df_print: paged
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
  base_dir: /var/lib/R/project-evalR
  calls_packages_path: /run/package-evals-traced.9/summarized-packages.fst
  corpus_path: /run/package-evals-traced.9/corpus.fst
---

The setup chunk has a variable `jan` with setting to run the script on
docker on Jan's mamchine. This should be removed.

@@

jan <- TRUE

knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) { # record the current time before each chunk
      now <<- Sys.time()
    } else { # calculate the time difference after a chunk
      res <- difftime(Sys.time(), now)
      # return a character string to show the time
      paste("Time for this code chunk to run:", round(res,2), units(res))
}}}))
now <- Sys.time()
knitr::opts_chunk$set(
  echo = TRUE,
  fig.retina = 2,
  fig.width = 10,
  cache.lazy = FALSE,
  time_it = TRUE
)

@@

```{r read,echo=F,cache=F}
read_fst(eval_calls_path) %>% as_tibble() -> E_raw
read_fst(corpus_path) %>% as_tibble() -> C_raw
#read_csv(paste0(base_dir, "/run/package-evals-traced.9/package-eval-static.csv")) -> P

E_raw %>% select(
  run_package = package,        # The package which was run to trigger the eval
  ev_variant = eval_function,   # Which eval variant was called
  duplicates = nb_ev_calls,     # Number of identical eval calls (weight)
  src_ref = eval_call_srcref,   # Source ref for the eval call site
  file,                         # Script file name (name of this run)
  ev_package =eval_source,      # Package of the call site
  arg_size = expr_expression_length, # Size in bytes of eval argument string
  opcodes = direct_interpreter_eval  # Non-recursive AST nodes traversed
) -> E                               # by the interpreter

C_raw -> C
corpus_size <- nrow(C)
```

@@

```{r nb-eval-call-sites, echo=F}
stopifnot(sum(is.na(E$src_ref))==0)  # No NAs in scr_ref!

@@

TBL <- overview_table(
# Sites triggered over the whole corpus:
r("nb eval call sites", sites) ,
# How many packages in the corpus have a single triggered call site:
r("nb one call site", sites_per_package %>% filter(n==1)),  
r("at least one call site", sites_per_package),
r("packages no calls to eval", corpus_size - nrow(sites_per_package)),
r("nb median call sites",  median(sites_per_package$n)),
r("max call sites", max_packages),
)
```

@@

```{r call-sites-violin-plot, echo=F}
E %>%
  group_by(ev_package) %>%
  mutate(sites=n_distinct(src_ref)) %>%
  ggplot(aes(x=ev_variant, y=sites, fill=ev_variant), na.rm=TRUE) +
  geom_violin(trim=FALSE, width=1.7) +
  # geom_jitter(shape=16, position=position_jitter(0.2)) +
  labs(y="Call sites per package") +
  theme(
    legend.position=c(0.5, 0.85),
    legend.box.background=element_rect(fill="white", size=0.1),
    legend.box="horizontal",
    axis.title.x=element_blank()
  )

@@

>>>>

---
title: "Usage metrics"
author: "Pierre Donat-Bouillud"
output:
    html_document:
        gallery: false
        toc: true
        toc_depth: 3
        toc_float: true
        df_print: paged
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
params:
  base_dir: /var/lib/R/project-evalR/run/package-evals-traced.10/
  calls_packages_path: summarized-packages.fst
  corpus_path: corpus.fst
  static_evals_path: package-evals-static.csv
---

The setup chunk has a variable `jan` with setting to run the script on
docker on Jan's machine. This should be removed.

@@

jan <- FALSE

knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) { # record the current time before each chunk
      now <<- Sys.time()
    } else { # calculate the time difference after a chunk
      res <- difftime(Sys.time(), now)
      # return a character string to show the time
      paste("Time for this code chunk to run:", round(res, 2), units(res))
    }
  }
}))
now <- Sys.time()
knitr::opts_chunk$set(
  echo = TRUE,
  fig.retina = 2,
  fig.width = 10,
  cache.lazy = FALSE,
  time_it = TRUE
)

@@

```{r read,echo=F,cache=F}
read_fst(eval_calls_path) %>% as_tibble() -> E_raw
read_fst(corpus_path) %>% as_tibble() -> C_raw

read_csv(eval_static_path) %>% semi_join(C_raw, by = "package") -> P
```

The two data sets are E and C for conciseness.  C is only used here to compute
the corpus size. E is used throughout. Recall that E_raw has the source data
with all columns. P corresponds to static information about the call sites.

```{r select-columns,echo=F}
E_raw %>% select(
  run_package = package, # The package which was run to trigger the eval
  ev_variant = eval_function, # Which eval variant was called
  duplicates = nb_ev_calls, # Number of identical eval calls (weight)
  src_ref = eval_call_srcref, # Source ref for the eval call site
  file, # Script file name (name of this run)
  ev_package = eval_source, # Package of the call site
  arg_size = expr_expression_length, # Size in bytes of eval argument string
  opcodes = direct_interpreter_eval # Non-recursive AST nodes traversed
) -> E # by the interpreter

C_raw -> C
corpus_size <- nrow(P)
```

@@

```{r nb-eval-call-sites, echo=F}
stopifnot(sum(is.na(E$src_ref)) == 0) # No NAs in scr_ref!

@@

TBL <- overview_table(
  # Sites triggered over the whole corpus:
  r("nb eval call sites", sites),
  # How many packages in the corpus have a single triggered call site:
  r("nb one call site", sites_per_package %>% filter(n == 1)),
  r("at least one call site", sites_per_package),
  r("packages no calls to eval", corpus_size - nrow(sites_per_package)),
  r("nb median call sites", median(sites_per_package$n)),
  r("max call sites", max_packages),
)
```

@@

```{r call-sites-violin-plot, echo=F}
E %>%
  group_by(ev_package) %>%
  mutate(sites = n_distinct(src_ref)) %>%
  ggplot(aes(x = ev_variant, y = sites, fill = ev_variant), na.rm = TRUE) +
  geom_violin(trim = FALSE, width = 1.7) +
  # geom_jitter(shape=16, position=position_jitter(0.2)) +
  labs(y = "Call sites per package") +
  theme(
    legend.position = c(0.5, 0.85),
    legend.box.background = element_rect(fill = "white", size = 0.1),
    legend.box = "horizontal",
    axis.title.x = element_blank()
  )

@@

How many call sites and how many packages did we exercise out of the total number of call sites of the corpus?

```{r exercised-call-sites}
static_sites <- nrow(P)

static_call_sites_per_package <- P %>%
  count(package, sort = TRUE)

static_at_least_one_call_site <- static_call_sites_per_package %>%
  filter(n >= 1)

r("nb static call sites", static_sites)
r("call sites percent", ratio(sites, static_sites))
r("nb static at least one call site", nrow(static_at_least_one_call_site))
r("static at least one call site percent", ratio(nrow(sites_per_package %>% filter(n >= 1)), nrow(static_at_least_one_call_site)))
```
